---
title: "Hw6"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(repos=structure(c(CRAN="YOUR FAVORITE MIRROR")))

install.packages("dplyr")
install.packages("tidyverse")
install.packages("ggplot2")
library(ggplot2)
library(tidyverse)
library(dplyr)
```



```{r}
crime <- read.delim("http://www.statsci.org/data/general/uscrime.txt")
data_scaled <- as.data.frame(scale(crime))

```


```{r}
PCA <- prcomp(crime[,1:15], scale=TRUE)
PCA
names(PCA)
```



```{r}
ggplot(as.data.frame(PCA$x), aes(x = PC1, y = PC2)) + geom_point()

```

```{r}
summary(PCA)
```

You can see that the first component accounts for 40% of variance. Rest of components have less importance as you can see from proportion of variance.We can conclude that this data set is governed by crime variable. 

```{r}
#Building linear regression model with first 5 terms
PC <- PCA$x[,1:5]
crimePC <- cbind(PC,crime[,16])
modelPCA <- lm(V6~., data = as.data.frame(crimePC))
summary(modelPCA)

```
Looking at the linear regression model, all terms except PC3 have a p value below 0.05. This shows 4 of the 5 terms are relevant factors. The adjusted R squared value is really low at 0.6 which shows the linear regression model doesn't fit as well from last weeks 8.2 solution.

Compared to my 8.2 solution this model is worse as the Rsquared was at least 0.7 so it was a much better fit. 




Below we are reconstructing model in terms of original variables.
```{r}
#b0 is our intercept
b0 <- modelPCA$coefficients[1]

#beta vector is created using the coefficients
betas <- modelPCA$coefficients[2:6]

b0

```
```{r}

#alpha vector is calculated below
alphas <- PCA$rotation[,1:5]%*%betas

#uscaling data below for beta and alpha and then calculating estimates
unscaledalpha <- alphas/sapply(crime[,1:15],sd)
betaunscaled <- b0 - sum(alphas*sapply(crime[,1:15],mean)/sapply(crime[,1:15],sd))

 est <- as.matrix(crime[,1:15]) %*% unscaledalpha + betaunscaled
 est
```
```{r}
SSE = sum((est - crime[,16])^2)
SStot = sum((crime[,16] - mean(crime[,16]))^2)
R2 <- 1-SSE/SStot


R2adjust <- R2 - (1-R2)*5/(nrow(crime)-5-1)
R2adjust
```

Looking at calculated R2 adjusted values you can see that the value is pretty decent at 0.6 which shows the model fit well but not as well as the linear regression from 8.2 which was higher at 0.7.

```{r}
a <- data.frame(M = 14.0,So = 0,Ed = 10.0, Po1 = 12.0,Po2 = 15.5,
                 LF = 0.640, M.F = 94.0,Pop = 150,NW = 1.1,U1 = 0.120,
                 U2 = 3.6, Wealth = 3200,Ineq = 20.1,Prob = 0.04, Time = 39.0)

pred_a <- data.frame(predict(PCA,a))

pred_model<-predict(modelPCA,pred_a)

pred_model
```

This prediction is within the range of the data set so it is a reasonble data point. From last weeks problem 8.2 we got a value 1304 and this is 1388 which is reasonably close to this value so the PCA mode gave close results. 